---
title: "AWS EMR Spot Instance Cluster for PGx Analysis"
author: "Jerome Dixon"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
editor: source
format:
  html:
    toc: true
    toc-depth: 4
    html-math-method: katex
    df-print: paged
    embed-resources: true
    fontsize: "11pt"
    code-fold: true
    code-summary: "Show the code"
    smooth-scroll: true
editor_options: 
  chunk_output_type: inline
execute:
  echo: true
  message: false
  warning: false
  eval: false
---

```{r sync-renv-bash}

library(jsonlite)
library(here)

cwd <- here::here()

Sys.setenv("CWD"=cwd)

```

### AWS EC2 Spot Advisor

-   This section provides a comparison between x2idn and x2iedn instance types. These are high-memory instances suitable for memory-intensive workloads. The table below shows key differences in specifications and use cases

| Feature | x2idn | x2iedn |
|------------------------|------------------------|------------------------|
| Architecture | x86_64 (Intel Cascade Lake) | Same |
| Memory per vCPU | \~8 GiB per vCPU | \~16 GiB per vCPU (2x more) |
| Memory (e.g. 16xlarge) | 512 GiB | 1,024 GiB |
| vCPUs (16xlarge) | 64 | 64 |
| Instance Storage | 1 x 1.9 TB NVMe SSD | 2 x 1.5 TB NVMe SSDs |
| Network Bandwidth | Up to 50 Gbps | Up to 100 Gbps |
| Use Case Focus | Cost-optimized memory | Max memory performance + bandwidth |
| ENA Express | No | Yes (lower latency, higher throughput) |
| Price | Lower | Higher (due to more memory, network, storage) |

-   Use case recommendations based on instance type characteristics:

Use x2idn if: - You're looking for a cost-effective high-memory instance - Your workload needs high memory but not the absolute maximum memory bandwidth - You're okay with 8 GiB/vCPU and lower network/storage

Use x2iedn if: - You need 1 TB+ of RAM and super-fast I/O - You're running in-memory databases (SAP HANA, Redis, etc.) - You're processing large model training, graph analytics, or real-time analytics - You want highest bandwidth & lowest latency networking

<https://aws.amazon.com/ec2/spot/instance-advisor/>

#### Spot Pricing

```{bash}

aws ec2 describe-spot-price-history \
  --instance-types x2idn.16xlarge x2idn.24xlarge \
  --product-descriptions "Linux/UNIX" \
  --start-time $(date -u +"%Y-%m-%dT%H:%M:%SZ") \
  --region us-east-1 \
  --query "SpotPriceHistory[*].[InstanceType,AvailabilityZone,SpotPrice]" \
  --output table

```

Note: No big price difference from 16xlarge and 24xlarge but 24xlarge is more available and therefore cheaper in a few sub-regions

-   Lowest price Instance Type

```{bash}

instance_type=$(aws ec2 describe-spot-price-history \
  --instance-types x2idn.16xlarge x2idn.24xlarge \
  --product-descriptions "Linux/UNIX" \
  --start-time $(date -u +"%Y-%m-%dT%H:%M:%SZ") \
  --region us-east-1 \
  --query "sort_by(SpotPriceHistory, &SpotPrice)[0].[InstanceType]" \
  --output text)
  
echo "$instance_type"
echo "$instance_type" > instance.txt

```

-   Lowest Price Availability Zone

```{bash}

availability_zone=$(aws ec2 describe-spot-price-history \
  --instance-types x2idn.16xlarge x2idn.24xlarge \
  --product-descriptions "Linux/UNIX" \
  --start-time $(date -u +"%Y-%m-%dT%H:%M:%SZ") \
  --region us-east-1 \
  --query "sort_by(SpotPriceHistory, &SpotPrice)[0].[AvailabilityZone]" \
  --output text)
  
echo "$availability_zone"
echo "$availability_zone" > zone.txt

```

-   Subnet for Lowest Price Instance

```{bash}

zone_id=$(cat zone.txt)

aws ec2 describe-subnets \
  --filters "Name=availability-zone,Values=us-east-1f" \
  --query 'Subnets[*].[SubnetId,AvailabilityZone]' \
  --output table

```

```{bash}

zone_id=$(cat zone.txt)

subnet_id=$(aws ec2 describe-subnets \
  --filters "Name=availability-zone,Values=$zone_id" \
  --query 'Subnets[*].[SubnetId]' \
  --output text)

echo "$subnet_id"
echo "$subnet_id" > subnet.txt

```

### x2idn.24xlarge EMR Cluster 

```{bash}

export aws_account=535362115856
subnet_id=$(cat subnet.txt)

aws emr create-cluster \
  --name "PGx-Analysis-2025" \
  --log-uri "s3://aws-logs-535362115856-us-east-1/elasticmapreduce" \
  --release-label "emr-7.5.0" \
  --service-role "arn:aws:iam::$aws_account:role/EMR_DefaultRole" \
  --security-configuration "pgx_emr2" \
  --managed-scaling-policy '{"ComputeLimits":{"UnitType":"Instances","MinimumCapacityUnits":2,"MaximumCapacityUnits":4,"MaximumOnDemandCapacityUnits":0,"MaximumCoreCapacityUnits":4}}' \
  --unhealthy-node-replacement \
  --ec2-attributes '{
    "InstanceProfile":"EMR_EC2_DefaultRole",
    "EmrManagedMasterSecurityGroup":"sg-0a5260eee9a20654e",
    "EmrManagedSlaveSecurityGroup":"sg-0f01580b24ea364e1",
    "ServiceAccessSecurityGroup":"sg-0e25102a082ef38d1",
    "KeyName":"mushin_pgx",
    "SubnetId":"$subnet_id"
  }' \
  --applications Name=Hadoop Name=Hive Name=JupyterEnterpriseGateway Name=Livy Name=Spark \
  --configurations '[{"Classification":"livy-conf","Properties":{"livy.server.session.timeout":"12h","livy.spark.deploy-mode":"cluster"}},{"Classification":"spark-hive-site","Properties":{"hive.metastore.client.factory.class":"com.amazonaws.glue.catalog.metastore.AWSGlueDataCatalogHiveClientFactory"}}]' \
  --instance-groups '[
    {
      "InstanceCount":1,
      "InstanceGroupType":"MASTER",
      "Market":"SPOT",
      "Name":"Primary",
      "InstanceType":"x2idn.24xlarge",
      "EbsConfiguration":{
        "EbsBlockDeviceConfigs":[
          {
            "VolumeSpecification":{
              "VolumeType":"gp2",
              "SizeInGB":192
            },
            "VolumesPerInstance":4
          }
        ]
      }
    },
    {
      "InstanceCount":1,
      "InstanceGroupType":"CORE",
      "Market":"SPOT",
      "Name":"Core",
      "InstanceType":"x2idn.24xlarge",
      "EbsConfiguration":{
        "EbsBlockDeviceConfigs":[
          {
            "VolumeSpecification":{
              "VolumeType":"gp2",
              "SizeInGB":192
            },
            "VolumesPerInstance":4
          }
        ]
      }
    }
  ]'\
  --bootstrap-actions '[{"Args":[],"Name":"pgx","Path":"s3://pgx-repository/emr-config/pgx_emr_studio_python39_bootstrap.sh"}]' \
  --auto-scaling-role "arn:aws:iam::$aws_account:role/EMR_AutoScaling_DefaultRole" \
  --scale-down-behavior "TERMINATE_AT_TASK_COMPLETION" \
  --ebs-root-volume-size "97" \
  --auto-termination-policy '{"IdleTimeout":3600}' \
  --region "us-east-1"

```

### Setup local environment 

For passing commands between R and Bash environments

```{r sync-renv-bash}

library(jsonlite)
library(here)

cwd <- here::here()

Sys.setenv("CWD"=cwd)

```

### Get EMR Cluster Info

```{sh get-emr-cluster-id}

aws emr list-clusters --active --profile pgx | tee ${CWD}/emr/cluster-config/cluster_info.json

```


```{r parse-emr-cluster-id}

cluster_json <- read_json(here("emr","cluster-config","cluster_info.json"))

cluster_id <- as.data.frame(cluster_json[["Clusters"]][[1]][["Id"]])

names(cluster_id)  <- "cluster_id"

write.table(cluster_id, here("emr","cluster-config","cluster_id.csv"), sep = ";", col.names = FALSE, row.names = FALSE)

cluster_id

# Save as environmental variable
Sys.setenv("CLUSTER_ID"=cluster_id)

```

#### Cluster Configuration

```{sh get-emr-cluster-config}

aws emr describe-cluster --cluster-id ${CLUSTER_ID} --profile pgx | tee ${CWD}/emr/cluster-config/cluster_description.json

```


```{r read-emr-cluster-config}

cluster_description_json <- read_json(here("emr","cluster-config","cluster_description.json"))

```

#### Cluster Head Node

```{sh get-emr-head-node}

aws emr list-instances --instance-group-type MASTER --cluster-id ${CLUSTER_ID} --profile pgx | tee ${CWD}/emr/cluster-config/emr_node.json

```

```{r parse-emr-head-node-info}

emr_json <- read_json(here("emr","cluster-config","emr_node.json"))

emr_dns <- emr_json[["Instances"]][[1]][["PublicDnsName"]]

emr_instance_id <- emr_json[["Instances"]][[1]][["Ec2InstanceId"]]

emr_instance_id <- as.data.frame(emr_instance_id)

write.table(emr_instance_id, here("emr","cluster-config","emr_instance_id.csv"), sep = ";", col.names = FALSE, row.names = FALSE)

emr_dns

emr_instance_id

Sys.setenv("EMR_NODE_ID"=emr_instance_id)

```

#### Allocate Elastic IP to AWS EMR Head Node

```{sh allocate-elastic-ip}

aws ec2 associate-address --instance-id ${EMR_NODE_ID} --allocation-id eipalloc-0d2b32dc1f19a9a4a --profile pgx

```

Setup your elastic IP or use EMR head node DNS. Don't forget to add :8787 at the end for RStudio port
Wait for cluster to build... {about 45 mins}

### Go to work!

```{r emr-rstudio-ip}

url <- "http://44.221.159.159:8787"

browseURL(url)

```

For failed cluster starts:

```{bash}

aws emr describe-cluster --cluster-id ${CLUSTER_ID} --profile pgx | tee ${CWD}/emr/cluster-config/bad_cluster_info.json

```

### EMR Studio

#### Permissions

```{bash}

aws emr list-studios --profile mushin


```


```{bash}

aws emr describe-studio --studio-id es-17EXJHZCNPQ7I1NFSVQNCOOPL --profile pgx

```


```{bash}

aws iam put-role-policy \
    --role-name AmazonEMRStudio_ServiceRole_1732182987289 \
    --policy-name AllowPassRoleForEMRExecution \
    --policy-document file://pass_role.json \
    --profile pgx

```


```{bash}

aws iam get-role --role-name AmazonEMRStudio_RuntimeRole_1732182987289 --profile pgx

```


```{bash}

aws iam get-role --role-name AmazonEMRStudio_RuntimeRole_1732182987289 --profile pgx

```


```{bash}

aws iam update-assume-role-policy \
    --role-name AmazonEMRStudio_RuntimeRole_1732182987289 \
    --policy-document file://trust_policy.json \
    --profile pgx

```


```{bash}

aws iam attach-role-policy \
    --role-name AmazonEMRStudio_RuntimeRole_1732182987289 \
    --policy-arn arn:aws:iam::aws:policy/service-role/AmazonElasticMapReduceEditorsRole \
    --profile pgx

```


```{bash}
aws iam put-role-policy \
    --role-name EMRStudio_User_Role \
    --policy-name AllowAddTagsOnEMRClusters \
    --policy-document file://add_tags_policy.json \
    --profile pgx

```

#### EMR Studio Cluster

```{bash}

aws emr create-cluster \
 --name "pgx-analysis-ascpt" \
 --log-uri "s3://athena-output-testing-pgx/elasticmapreduce" \
 --release-label "emr-7.5.0" \
 --service-role "EMR_DefaultRole" \
 --unhealthy-node-replacement \
 --ec2-attributes '{"InstanceProfile":"EMR_EC2_DefaultRole","EmrManagedMasterSecurityGroup":"sg-0a5260eee9a20654e","EmrManagedSlaveSecurityGroup":"sg-0f01580b24ea364e1","ServiceAccessSecurityGroup":"sg-0e25102a082ef38d1","SubnetId":"subnet-03f689fd2cb8cbb2c"}' \
 --tags 'creator=CREATE_CLUSTER_TAB_IN_JUPYTERLAB_PLUGIN' \
 --applications Name=Spark Name=Hadoop Name=Livy Name=JupyterEnterpriseGateway \
 --instance-groups '[{"InstanceCount":1,"InstanceGroupType":"MASTER","Name":"master","InstanceType":"m5.xlarge","EbsConfiguration":{"EbsBlockDeviceConfigs":[{"VolumeSpecification":{"VolumeType":"gp2","SizeInGB":32},"VolumesPerInstance":2}]}},{"InstanceCount":2,"InstanceGroupType":"CORE","Name":"slave","InstanceType":"m5.xlarge","EbsConfiguration":{"EbsBlockDeviceConfigs":[{"VolumeSpecification":{"VolumeType":"gp2","SizeInGB":32},"VolumesPerInstance":2}]}}]' \
 --scale-down-behavior "TERMINATE_AT_TASK_COMPLETION" \
 --bootstrap-actions '[{"Args":[],"Name":"pgx","Path":"s3://pgx-repository/emr-config/pgx_emr_studio_python39_bootstrap.sh"}]' \
 --auto-termination-policy '{"IdleTimeout":3600}' \
 --region "us-east-1" \
 --profile pgx
```


