---
title: "EMR Spot Cluster for PGx Analysis - Wide"
author: "Jerome Dixon"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
editor: source
format:
  html:
    toc: true
    toc-depth: 5
    html-math-method: katex
    df-print: paged
    embed-resources: true
    fontsize: "11pt"
    code-fold: true
    code-summary: "Show the code"
    smooth-scroll: true
editor_options: 
  chunk_output_type: inline
execute:
  echo: true
  message: false
  warning: false
  eval: false
---

```{r sync-renv-bash}

library(jsonlite)
library(here)

cwd <- here::here()

Sys.setenv("CWD"=cwd)

```

### AWS EC2 Spot Advisor

-   # Updated Instance Type Comparison for Memory-Intensive Workloads

This section provides a comparison between x2idn, x2iedn, and x2gd instance types. These are high-memory instances suitable for memory-intensive workloads. The table below shows key differences in specifications and use cases:

| Feature | x2idn | x2iedn | x2gd |
|------------------------|------------------------|------------------------|------------------------|
| Architecture | x86_64 (Intel Cascade Lake) | Same | ARM (AWS Graviton2) |
| Memory per vCPU | ~8 GiB per vCPU | ~16 GiB per vCPU (2x more) | ~16 GiB per vCPU |
| Memory (2xlarge) | 128 GiB | 256 GiB | 128 GiB |
| vCPUs (2xlarge) | 8 | 8 | 8 |
| Instance Storage | NVMe SSD | Larger NVMe SSDs | Local NVMe SSD (474 GB) |
| Network Bandwidth | Up to 25 Gbps | Higher (Up to 25 Gbps) | Up to 10 Gbps |
| Use Case Focus | Cost-optimized memory | Max memory performance + bandwidth | Best price-performance for memory workloads |
| ARM Compatibility | No | No | Required (Graviton2) |
| Price-Performance | Good | Premium | 20-40% better than x86 equivalents |
| ENA Express | No | Yes (lower latency, higher throughput) | No |

## Use case recommendations based on instance type characteristics:

**Use x2idn if:**
- You're looking for a cost-effective high-memory instance with x86 architecture
- Your workload needs high memory but not the absolute maximum memory bandwidth
- You're okay with 8 GiB/vCPU and need good network/storage performance

**Use x2iedn if:**
- You need maximum RAM and super-fast I/O on x86 architecture
- You're running in-memory databases (SAP HANA, Redis, etc.)
- You're processing large model training, graph analytics, or real-time analytics
- You want highest bandwidth & lowest latency networking

**Use x2gd if:**
- Your application is ARM-compatible
- You want the best price-performance ratio for memory-intensive workloads
- You need a good memory-to-CPU ratio (16 GiB/vCPU)
- Your Spark/database workloads are cost-sensitive but memory-hungry
- You can trade some network bandwidth for cost savings

The x2gd instances offer an excellent balance of performance and cost for compatible workloads, particularly for data processing pipelines that don't require maximum network throughput.

<https://aws.amazon.com/ec2/spot/instance-advisor/>

```{r}

browseURL("https://aws.amazon.com/ec2/spot/instance-advisor/")

```


### Spot Pricing

```{bash}

aws ec2 describe-spot-price-history \
  --instance-types x2iedn.2xlarge x2iedn.xlarge \
  --product-descriptions "Linux/UNIX" \
  --start-time $(date -u +"%Y-%m-%dT%H:%M:%SZ") \
  --region us-east-1 \
  --query "SpotPriceHistory[*].[InstanceType,AvailabilityZone,SpotPrice]" \
  --output table

```

Note: x2iedn.xlarge is preferred 

#### Lowest price Instance Type

```{bash}

instance_type=$(aws ec2 describe-spot-price-history \
  --instance-types x2iedn.xlarge x2iedn.2xlarge \
  --product-descriptions "Linux/UNIX" \
  --start-time $(date -u +"%Y-%m-%dT%H:%M:%SZ") \
  --region us-east-1 \
  --query "sort_by(SpotPriceHistory, &SpotPrice)[0].[InstanceType]" \
  --output text)
  
echo "$instance_type"
echo "$instance_type" > instance.txt

```

#### Lowest Price Availability Zone

```{bash}

instance_id=$(cat instance.txt)

availability_zone=$(aws ec2 describe-spot-price-history \
  --instance-types $instance_id \
  --product-descriptions "Linux/UNIX" \
  --start-time $(date -u +"%Y-%m-%dT%H:%M:%SZ") \
  --region us-east-1 \
  --query "sort_by(SpotPriceHistory, &SpotPrice)[0].[AvailabilityZone]" \
  --output text)
  
echo "$availability_zone"
echo "$availability_zone" > zone.txt

```

-   Subnet for Lowest Price Instance in EMR VPC

```{bash}

zone_id=$(cat zone.txt)
emr_vpc_id="vpc-0ea3d9b6c6c64d0af"
  
aws ec2 describe-subnets \
  --filters "Name=availability-zone,Values=$zone_id" "Name=vpc-id,Values=$emr_vpc_id" \
  --query 'Subnets[*].[SubnetId,AvailabilityZone,VpcId]' \
  --output table


```

#### If Subnet Available in EMR VPC
(if not available - emr_vpc_config.qmd)

```{bash}

zone_id=$(cat zone.txt)
emr_vpc_id="vpc-0ea3d9b6c6c64d0af"
  
vpc_subnet_id=$(aws ec2 describe-subnets \
  --filters "Name=availability-zone,Values=$zone_id" "Name=vpc-id,Values=$emr_vpc_id" \
  --query 'Subnets[*].[SubnetId]' \
  --output text)
  
echo "$vpc_subnet_id"
echo "$vpc_subnet_id" > vpc_subnet.txt

```


### x2iedn.xlarge EMR Cluster 

```{bash}

instance_type=$(cat instance.txt)

cat > instance-fleet.json <<EOF
[
  {
    "Name": "MasterFleet",
    "InstanceFleetType": "MASTER",
    "TargetOnDemandCapacity": 1,
    "TargetSpotCapacity": 0,
    "InstanceTypeConfigs": [
      {
        "InstanceType": "$instance_type",
        "EbsConfiguration": {
          "EbsBlockDeviceConfigs": [
            {
              "VolumeSpecification": {
                "VolumeType": "gp2",
                "SizeInGB": 192
              },
              "VolumesPerInstance": 4
            }
          ]
        }
      }
    ]
  },
  {
    "Name": "CoreFleet",
    "InstanceFleetType": "CORE",
    "TargetOnDemandCapacity": 25,
    "TargetSpotCapacity": 75,
    "InstanceTypeConfigs": [
      {
        "InstanceType": "$instance_type",
        "WeightedCapacity": 1,
        "EbsConfiguration": {
            "EbsBlockDeviceConfigs": [
              {
                "VolumeSpecification": {
                  "VolumeType": "gp3",
                  "SizeInGB": 256,
                  "Throughput": 250,
                  "Iops": 3000
                },
                "VolumesPerInstance": 1
              }
            ]
         }
      },
      {
        "InstanceType": "r5b.2xlarge",
        "WeightedCapacity": 8,
        "EbsConfiguration": {
            "EbsBlockDeviceConfigs": [
              {
                "VolumeSpecification": {
                  "VolumeType": "gp3",
                  "SizeInGB": 256,
                  "Throughput": 250,
                  "Iops": 3000
                },
                "VolumesPerInstance": 1
              }
            ]
          }
      },
      {
        "InstanceType": "r6id.xlarge",
        "WeightedCapacity": 1,
        "EbsConfiguration": {
            "EbsBlockDeviceConfigs": [
              {
                "VolumeSpecification": {
                  "VolumeType": "gp3",
                  "SizeInGB": 256,
                  "Throughput": 250,
                  "Iops": 3000
                },
                "VolumesPerInstance": 1
              }
            ]
         }
      }
    ],
    "LaunchSpecifications": {
      "SpotSpecification": {
        "TimeoutDurationMinutes": 10,
        "TimeoutAction": "SWITCH_TO_ON_DEMAND",
        "AllocationStrategy": "capacity-optimized"
      }
    }
  }
]
EOF

```

```{bash}

export aws_account=535362115856
subnet_id=$(cat vpc_subnet.txt)


ec2_attrs=$(cat <<EOF
{
  "InstanceProfile": "EMR_EC2_DefaultRole",
  "EmrManagedMasterSecurityGroup": "sg-0a5260eee9a20654e",
  "EmrManagedSlaveSecurityGroup": "sg-0f01580b24ea364e1",
  "ServiceAccessSecurityGroup": "sg-0e25102a082ef38d1",
  "KeyName": "mushin_pgx",
  "SubnetIds": [
  "subnet-0c0e86106b994e07c",
  "subnet-0a8809358f7fef905",
  "subnet-0802ec666c4a90f0d",
  "subnet-03f689fd2cb8cbb2c"
]

}
EOF
)


aws emr create-cluster \
  --name "PGx-Analysis-2025-wide" \
  --log-uri "s3://aws-logs-535362115856-us-east-1/elasticmapreduce" \
  --release-label "emr-7.5.0" \
  --service-role "arn:aws:iam::$aws_account:role/EMR_DefaultRole" \
  --security-configuration "pgx_emr2" \
  --managed-scaling-policy '{"ComputeLimits":{"UnitType":"InstanceFleetUnits","MinimumCapacityUnits":25,"MaximumCapacityUnits":120,"MaximumOnDemandCapacityUnits":45,"MaximumCoreCapacityUnits":120}}' \
  --unhealthy-node-replacement \
  --ec2-attributes "$ec2_attrs" \
  --applications Name=Hadoop Name=Hive Name=JupyterEnterpriseGateway Name=Livy Name=Spark \
  --configurations file://config/emr_config.json \
  --instance-fleets file://instance-fleet.json \
  --bootstrap-actions '[{"Args":[],"Name":"pgx","Path":"s3://pgx-repository/emr-config/emr_bootstrap.sh"}]' \
  --scale-down-behavior "TERMINATE_AT_TASK_COMPLETION" \
  --ebs-root-volume-size "97" \
  --auto-termination-policy '{"IdleTimeout":3600}' \
  --region "us-east-1"

```


```{bash}

aws emr create-cluster \
 --name "pgx-analysis-X2iedn-xlarge" \
 --log-uri "s3://athena-output-testing-pgx/elasticmapreduce" \
 --release-label "emr-7.5.0" \
 --service-role "arn:aws:iam::535362115856:role/EMR_DefaultRole" \
 --security-configuration "pgx_emr2" \
 --managed-scaling-policy '{"ComputeLimits":{"UnitType":"InstanceFleetUnits","MinimumCapacityUnits":3,"MaximumCapacityUnits":120,"MaximumOnDemandCapacityUnits":45,"MaximumCoreCapacityUnits":120}}' \
 --unhealthy-node-replacement \
 --ec2-attributes '{"InstanceProfile":"EMR_EC2_DefaultRole","EmrManagedMasterSecurityGroup":"sg-0a5260eee9a20654e","EmrManagedSlaveSecurityGroup":"sg-0f01580b24ea364e1","KeyName":"mushin_pgx","ServiceAccessSecurityGroup":"sg-0e25102a082ef38d1","SubnetIds":["subnet-03f689fd2cb8cbb2c"]}' \
 --applications Name=Hadoop Name=JupyterEnterpriseGateway Name=Livy Name=Spark \
 --configurations file://config/emr_config.json \
 --instance-fleets '[{"Name":"Core","InstanceFleetType":"CORE","TargetSpotCapacity":0,"TargetOnDemandCapacity":3,"LaunchSpecifications":{"SpotSpecification":{"TimeoutDurationMinutes":60,"TimeoutAction":"TERMINATE_CLUSTER","AllocationStrategy":"PRICE_CAPACITY_OPTIMIZED"},"OnDemandSpecification":{"AllocationStrategy":"LOWEST_PRICE"}},"InstanceTypeConfigs":[{"WeightedCapacity":4,"EbsConfiguration":{},"BidPriceAsPercentageOfOnDemandPrice":100,"InstanceType":"x2iedn.xlarge"}]},{"Name":"Primary","InstanceFleetType":"MASTER","TargetSpotCapacity":0,"TargetOnDemandCapacity":1,"LaunchSpecifications":{"OnDemandSpecification":{"AllocationStrategy":"LOWEST_PRICE"}},"InstanceTypeConfigs":[{"WeightedCapacity":1,"EbsConfiguration":{},"BidPriceAsPercentageOfOnDemandPrice":100,"InstanceType":"x2iedn.xlarge"}]}]' \
 --bootstrap-actions '[{"Args":[],"Name":"pgx","Path":"s3://pgx-repository/emr-config/emr_bootstrap.sh"}]' \
 --scale-down-behavior "TERMINATE_AT_TASK_COMPLETION" \
 --ebs-root-volume-size "97" \
 --ebs-root-volume-iops "3000" \
 --auto-termination-policy '{"IdleTimeout":1800}' \
 --region "us-east-1"
 
 
```



#### EMR Wide Configuration

```{python eval=FALSE}


from pyspark.sql import SparkSession
from pyspark.sql import functions as F
from pyspark.sql.types import IntegerType, DoubleType
from pyspark.sql.functions import col, max as spark_max, min as spark_min

spark = SparkSession.builder \
    .appName("PGx_Cohort_Dataset_Builder") \
    .config("spark.master", "yarn") \
    .config("spark.executor.memory", "16g") \
    .config("spark.driver.memory", "8g") \
    .config("spark.executor.memoryOverhead", "6g") \
    .config("spark.executor.cores", "4") \
    .config("spark.dynamicAllocation.enabled", "true") \
    .config("spark.dynamicAllocation.initialExecutors", "10") \
    .config("spark.dynamicAllocation.minExecutors", "2") \
    .config("spark.dynamicAllocation.maxExecutors", "200") \
    .config("spark.dynamicAllocation.executorIdleTimeout", "30s") \
    .config{"spark.dynamicAllocation.schedulerBacklogTimeout", "0.5s") \
    .config("spark.jars.packages.resolve.transitive", "true") \
    .config("spark.executor.extraJavaOptions", "--add-exports java.base/sun.net.util=ALL-UNNAMED") \
    .config("spark.network.timeout", "1200s") \
    .config("spark.rpc.askTimeout", "1200s") \
    .config("spark.sql.adaptive.skewJoin.enabled", "true") \
    .getOrCreate()

```