---
title: "EMR GPU Cluster for PGx Analysis on EMR Studio"
author: "Jerome Dixon"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
editor: source
format:
  html:
    toc: true
    toc-depth: 5
    html-math-method: katex
    df-print: paged
    embed-resources: true
    fontsize: "11pt"
    code-fold: true
    code-summary: "Show the code"
    smooth-scroll: true
editor_options: 
  chunk_output_type: inline
execute:
  echo: true
  message: false
  warning: false
  eval: false
---

```{r sync-renv-bash}

library(jsonlite)
library(here)

cwd <- here::here()

Sys.setenv("CWD"=cwd)

```

### AWS EC2 Spot Advisor

-   # Updated Instance Type Comparison for Memory-Intensive Workloads

This section provides a comparison between x2idn, x2iedn, and x2gd instance types. These are high-memory instances suitable for memory-intensive workloads. The table below shows key differences in specifications and use cases:

| Feature | x2idn | x2iedn | x2gd |
|------------------------|------------------------|------------------------|------------------------|
| Architecture | x86_64 (Intel Cascade Lake) | Same | ARM (AWS Graviton2) |
| Memory per vCPU | ~8 GiB per vCPU | ~16 GiB per vCPU (2x more) | ~16 GiB per vCPU |
| Memory (2xlarge) | 128 GiB | 256 GiB | 128 GiB |
| vCPUs (2xlarge) | 8 | 8 | 8 |
| Instance Storage | NVMe SSD | Larger NVMe SSDs | Local NVMe SSD (474 GB) |
| Network Bandwidth | Up to 25 Gbps | Higher (Up to 25 Gbps) | Up to 10 Gbps |
| Use Case Focus | Cost-optimized memory | Max memory performance + bandwidth | Best price-performance for memory workloads |
| ARM Compatibility | No | No | Required (Graviton2) |
| Price-Performance | Good | Premium | 20-40% better than x86 equivalents |
| ENA Express | No | Yes (lower latency, higher throughput) | No |

## Use case recommendations based on instance type characteristics:

**Use x2idn if:**
- You're looking for a cost-effective high-memory instance with x86 architecture
- Your workload needs high memory but not the absolute maximum memory bandwidth
- You're okay with 8 GiB/vCPU and need good network/storage performance

**Use x2iedn if:**
- You need maximum RAM and super-fast I/O on x86 architecture
- You're running in-memory databases (SAP HANA, Redis, etc.)
- You're processing large model training, graph analytics, or real-time analytics
- You want highest bandwidth & lowest latency networking

**Use x2gd if:**
- Your application is ARM-compatible
- You want the best price-performance ratio for memory-intensive workloads
- You need a good memory-to-CPU ratio (16 GiB/vCPU)
- Your Spark/database workloads are cost-sensitive but memory-hungry
- You can trade some network bandwidth for cost savings

The x2gd instances offer an excellent balance of performance and cost for compatible workloads, particularly for data processing pipelines that don't require maximum network throughput.

<https://aws.amazon.com/ec2/spot/instance-advisor/>

```{r}

browseURL("https://aws.amazon.com/ec2/spot/instance-advisor/")

```
### Spot Pricing

```{bash}

aws ec2 describe-spot-price-history \
  --instance-types x2idn.16xlarge x2idn.24xlarge \
  --product-descriptions "Linux/UNIX" \
  --start-time $(date -u +"%Y-%m-%dT%H:%M:%SZ") \
  --region us-east-1 \
  --query "SpotPriceHistory[*].[InstanceType,AvailabilityZone,SpotPrice]" \
  --output table

```

Note: No big price difference from 16xlarge and 24xlarge but 24xlarge is more available and therefore cheaper in a few sub-regions

-   Lowest price Instance Type

```{bash}

instance_type=$(aws ec2 describe-spot-price-history \
  --instance-types x2idn.16xlarge x2idn.24xlarge \
  --product-descriptions "Linux/UNIX" \
  --start-time $(date -u +"%Y-%m-%dT%H:%M:%SZ") \
  --region us-east-1 \
  --query "sort_by(SpotPriceHistory, &SpotPrice)[0].[InstanceType]" \
  --output text)
  
echo "$instance_type"
echo "$instance_type" > config/instance.txt

```

-   Lowest Price Availability Zone

```{bash}

availability_zone=$(aws ec2 describe-spot-price-history \
  --instance-types x2idn.16xlarge x2idn.24xlarge \
  --product-descriptions "Linux/UNIX" \
  --start-time $(date -u +"%Y-%m-%dT%H:%M:%SZ") \
  --region us-east-1 \
  --query "sort_by(SpotPriceHistory, &SpotPrice)[0].[AvailabilityZone]" \
  --output text)
  
echo "$availability_zone"
echo "$availability_zone" > config/zone.txt

```

-   Subnet for Lowest Price Instance in EMR VPC

```{bash}

zone_id=$(cat config/zone.txt)
emr_vpc_id="vpc-0ea3d9b6c6c64d0af"
  
aws ec2 describe-subnets \
  --filters "Name=availability-zone,Values=$zone_id" "Name=vpc-id,Values=$emr_vpc_id" \
  --query 'Subnets[*].[SubnetId,AvailabilityZone,VpcId]' \
  --output table


```

#### If Subnet Available in EMR VPC

```{bash}

zone_id=$(cat config/zone.txt)
emr_vpc_id="vpc-0ea3d9b6c6c64d0af"
  
vpc_subnet_id=$(aws ec2 describe-subnets \
  --filters "Name=availability-zone,Values=$zone_id" "Name=vpc-id,Values=$emr_vpc_id" \
  --query 'Subnets[*].[SubnetId]' \
  --output text)
  
echo "$vpc_subnet_id"
echo "$vpc_subnet_id" > config/vpc_subnet.txt

```

#### If Subnet Not Available

##### Describe EMR VPC

```{bash}

vpc_id="vpc-0ea3d9b6c6c64d0af"

aws ec2 describe-vpcs \
  --vpc-ids "$vpc_id" \
  --query 'Vpcs[*].[VpcId,CidrBlock,State,IsDefault]' \
  --output table

```
`

##### Add New Subnet

```{bash eval=FALSE}

vpc_id="vpc-0ea3d9b6c6c64d0af"

aws ec2 create-subnet \
  --vpc-id "$vpc_id" \
  --cidr-block 172.31.160.0/20 \
  --availability-zone us-east-1f \
  --tag-specifications 'ResourceType=subnet,Tags=[{Key=Name,Value=emr-studio-private-us-east-1f}]'


```


```{bash}

aws ec2 describe-subnets \
  --filters Name=vpc-id,Values=vpc-0ea3d9b6c6c64d0af \
           Name=tag:Name,Values=emr-studio-private-us-east-1f \
  --query "Subnets[*].SubnetId" \
  --output text > config/vpc_subnet.txt

```

- Ensure private

```{bash}

subnet_id=$(cat config/vpc_subnet.txt)

aws ec2 modify-subnet-attribute \
  --subnet-id "$subnet_id" \
  --no-map-public-ip-on-launch


```

##### Link Subnet to Routing Table

- Must use NAT Gateway

```{bash}

vpc_id="vpc-0ea3d9b6c6c64d0af"
subnet_id=$(cat config/new_subnet.txt)

aws ec2 describe-route-tables \
  --filters Name=vpc-id,Values="$vpc_id" \
            Name=route.nat-gateway-id,Values=* \
  --query 'RouteTables[*].{RouteTableId:RouteTableId,NatGatewayId:Routes[?NatGatewayId]|[0].NatGatewayId}' \
  --output table


```


```{bash}

aws ec2 describe-route-tables \
  --filters Name=route.nat-gateway-id,Values=* \
  --query 'RouteTables[0].RouteTableId' \
  --output text > config/route_table.txt


```

- Update to use Routing Table with NAT Gateway

```{bash}

subnet_id=$(cat config/new_subnet.txt)
route_table_id=$(cat config/route_table.txt)

aws ec2 associate-route-table \
  --subnet-id "$subnet_id" \
  --route-table-id "$route_table_id"


```


```{bash}

zone_id=$(cat config/zone.txt)

subnet_id=$(aws ec2 describe-subnets \
  --filters "Name=availability-zone,Values=$zone_id" \
  --query 'Subnets[*].[SubnetId]' \
  --output text)

echo "$subnet_id"
echo "$subnet_id" > config/vpc_subnet.txt

```

### x2idn.24xlarge EMR Cluster 

```{bash}

cat > config/instance-fleets.json <<EOF
[
  {
    "Name": "MasterFleet",
    "InstanceFleetType": "MASTER",
    "TargetOnDemandCapacity": 0,
    "TargetSpotCapacity": 1,
    "InstanceTypeConfigs": [
      {
        "InstanceType": "x2idn.24xlarge",
        "EbsConfiguration": {
          "EbsBlockDeviceConfigs": [
            {
              "VolumeSpecification": {
                "VolumeType": "gp2",
                "SizeInGB": 192
              },
              "VolumesPerInstance": 4
            }
          ]
        }
      }
    ]
  },
  {
    "Name": "CoreFleet",
    "InstanceFleetType": "CORE",
    "TargetOnDemandCapacity": 0,
    "TargetSpotCapacity": 1,
    "InstanceTypeConfigs": [
      {
        "InstanceType": "x2idn.24xlarge",
        "EbsConfiguration": {
          "EbsBlockDeviceConfigs": [
            {
              "VolumeSpecification": {
                "VolumeType": "gp2",
                "SizeInGB": 192
              },
              "VolumesPerInstance": 4
            }
          ]
        }
      }
    ]
  }
]

EOF

```


```{bash}

export aws_account=535362115856
subnet_id=$(cat config/vpc_subnet.txt)


ec2_attrs=$(cat <<EOF
{
  "InstanceProfile": "EMR_EC2_DefaultRole",
  "EmrManagedMasterSecurityGroup": "sg-0a5260eee9a20654e",
  "EmrManagedSlaveSecurityGroup": "sg-0f01580b24ea364e1",
  "ServiceAccessSecurityGroup": "sg-0e25102a082ef38d1",
  "KeyName": "mushin_pgx",
  "SubnetId": "$subnet_id"
}
EOF
)


aws emr create-cluster \
  --name "PGx-Analysis-2025" \
  --log-uri "s3://aws-logs-535362115856-us-east-1/elasticmapreduce" \
  --release-label "emr-7.5.0" \
  --service-role "arn:aws:iam::$aws_account:role/EMR_DefaultRole" \
  --security-configuration "pgx_emr2" \
  --managed-scaling-policy '{"ComputeLimits":{"UnitType":"InstanceFleetUnits","MinimumCapacityUnits":2,"MaximumCapacityUnits":4,"MaximumOnDemandCapacityUnits":0,"MaximumCoreCapacityUnits":4}}' \
  --unhealthy-node-replacement \
  --ec2-attributes "$ec2_attrs" \
  --applications Name=Hadoop Name=Hive Name=JupyterEnterpriseGateway Name=Livy Name=Spark \
  --configurations file://config/emr_config_wide.json \
  --instance-fleets file://config/instance-fleets.json \
  --bootstrap-actions '[{"Args":[],"Name":"pgx","Path":"s3://pgx-repository/emr-config/pgx_emr_studio_python39_bootstrap.sh"}]' \
  --scale-down-behavior "TERMINATE_AT_TASK_COMPLETION" \
  --ebs-root-volume-size "97" \
  --auto-termination-policy '{"IdleTimeout":3600}' \
  --region "us-east-1"

```
