---
title: "EMR Cluster"
author: "Jerome Dixon"
date: December 27, 2023
editor: source
format:
  html:
    toc: true
    toc-depth: 4
    html-math-method: katex
    df-print: paged
    embed-resources: true
    fontsize: "11pt"
    code-fold: true
    code-summary: "Show the code"
    smooth-scroll: true
editor_options: 
  chunk_output_type: inline
execute:
  echo: true
  message: false
  warning: false
  eval: false
---

### GPU (PySpark boostrap)

```{bash create-emr-cluster-gpu}

aws emr create-cluster \
 --name "PGx_Py_GPU" \
 --log-uri "s3n://aws-logs-$aws_account-us-east-1/elasticmapreduce/" \
 --release-label "emr-6.15.0" \
 --service-role "arn:aws:iam::$aws_account:role/EMR_DefaultRole" \
 --managed-scaling-policy '{"ComputeLimits":{"UnitType":"Instances","MinimumCapacityUnits":2,"MaximumCapacityUnits":4,"MaximumOnDemandCapacityUnits":3,"MaximumCoreCapacityUnits":3}}' \
 --ec2-attributes '{"InstanceProfile":"EMR_EC2_DefaultRole","EmrManagedMasterSecurityGroup":"sg-063a1a27f01fac9b1","EmrManagedSlaveSecurityGroup":"sg-040af67bc708a5425","SubnetId":"subnet-07950d9f2eb4d936d"}' \
 --applications Name=Hadoop Name=Spark \
 --configurations '[{"Classification":"spark","Properties":{"enableSparkRapids":"true"}},{"Classification":"yarn-site","Properties":{"yarn.nodemanager.container-executor.class":"org.apache.hadoop.yarn.server.nodemanager.LinuxContainerExecutor","yarn.nodemanager.linux-container-executor.cgroups.hierarchy":"yarn","yarn.nodemanager.linux-container-executor.cgroups.mount":"true","yarn.nodemanager.linux-container-executor.cgroups.mount-path":"/sys/fs/cgroup","yarn.nodemanager.resource-plugins":"yarn.io/gpu","yarn.nodemanager.resource-plugins.gpu.allowed-gpu-devices":"auto","yarn.nodemanager.resource-plugins.gpu.path-to-discovery-executables":"/usr/bin","yarn.resource-types":"yarn.io/gpu"}},{"Classification":"container-executor","Configurations":[{"Classification":"gpu","Properties":{"module.enabled":"true"}},{"Classification":"cgroups","Properties":{"root":"/sys/fs/cgroup","yarn-hierarchy":"yarn"}}],"Properties":{}},{"Classification":"spark-defaults","Properties":{"spark.executor.cores":"2","spark.executor.extraLibraryPath":"/usr/local/cuda/targets/x86_64-linux/lib:/usr/local/cuda/extras/CUPTI/lib64:/usr/local/cuda/compat/lib:/usr/local/cuda/lib:/usr/local/cuda/lib64:/usr/lib/hadoop/lib/native:/usr/lib/hadoop-lzo/lib/native:/docker/usr/lib/hadoop/lib/native:/docker/usr/lib/hadoop-lzo/lib/native","spark.executor.memoryOverhead":"2G","spark.executor.resource.gpu.amount":"1","spark.executor.resource.gpu.discoveryScript":"/usr/lib/spark/scripts/gpu/getGpusResources.sh","spark.locality.wait":"0s","spark.plugins":"com.nvidia.spark.SQLPlugin","spark.rapids.memory.pinnedPool.size":"0","spark.rapids.sql.concurrentGpuTasks":"1","spark.sql.files.maxPartitionBytes":"512m","spark.sql.shuffle.partitions":"200","spark.submit.pyFiles":"/usr/lib/spark/jars/xgboost4j-spark_3.0-1.4.2-0.3.0.jar","spark.task.cpus":"1","spark.task.resource.gpu.amount":"0.5"}},{"Classification":"capacity-scheduler","Properties":{"yarn.scheduler.capacity.resource-calculator":"org.apache.hadoop.yarn.util.resource.DominantResourceCalculator"}},{"Classification":"spark-hive-site","Properties":{"hive.metastore.client.factory.class":"com.amazonaws.glue.catalog.metastore.AWSGlueDataCatalogHiveClientFactory"}}]' \
 --instance-groups '[{"InstanceCount":1,"InstanceGroupType":"CORE","Name":"CORE","InstanceType":"g4dn.2xlarge","EbsConfiguration":{"EbsBlockDeviceConfigs":[]}},{"InstanceCount":1,"InstanceGroupType":"MASTER","Name":"MASTER","InstanceType":"m5.16xlarge","EbsConfiguration":{"EbsBlockDeviceConfigs":[{"VolumeSpecification":{"VolumeType":"gp2","SizeInGB":256},"VolumesPerInstance":4}]}}]' \
 --bootstrap-actions '[{"Args":[],"Name":"emr_config","Path":"s3://pgx-terraform/emr-config/pgx_emr_gpu_bootstrap.sh"}]' \
 --scale-down-behavior "TERMINATE_AT_TASK_COMPLETION" \
 --ebs-root-volume-size "97" \
 --region "us-east-1" \
 --profile pgx

```

## M5.16XLarge EMR Cluster (PySpark bootsrap) 

CatBoost not optimized for GPU. Using (1) m5.16xlarge Driver Node with (10) m5.16xlarge Core Node (CatBoost memory intensive)

```{bash}

aws emr create-cluster \
 --name "PGx_PySpark" \
 --log-uri "s3n://aws-logs-$aws_account-us-east-1/elasticmapreduce/" \
 --release-label "emr-6.15.0" \
 --service-role "arn:aws:iam::$aws_account:role/EMR_DefaultRole" \
 --security-configuration "pgx_emr2" \
 --managed-scaling-policy '{"ComputeLimits":{"UnitType":"Instances","MinimumCapacityUnits":2,"MaximumCapacityUnits":20,"MaximumOnDemandCapacityUnits":20,"MaximumCoreCapacityUnits":20}}' \
 --termination-protected \
 --ec2-attributes '{"InstanceProfile":"EMR_EC2_DefaultRole","EmrManagedMasterSecurityGroup":"sg-063a1a27f01fac9b1","EmrManagedSlaveSecurityGroup":"sg-040af67bc708a5425","AdditionalMasterSecurityGroups":[],"AdditionalSlaveSecurityGroups":[],"SubnetId":"subnet-07950d9f2eb4d936d"}' \
 --applications Name=Ganglia Name=Hadoop Name=Hive Name=Spark \
 --configurations '[{"Classification":"spark-hive-site","Properties":{"hive.metastore.client.factory.class":"com.amazonaws.glue.catalog.metastore.AWSGlueDataCatalogHiveClientFactory"}},{"Classification":"hive-site","Properties":{"hive.metastore.client.factory.class":"com.amazonaws.glue.catalog.metastore.AWSGlueDataCatalogHiveClientFactory"}}]' \
 --instance-groups '[{"InstanceCount":10,"InstanceGroupType":"CORE","Name":"Core","InstanceType":"m5.16xlarge","EbsConfiguration":{"EbsBlockDeviceConfigs":[{"VolumeSpecification":{"VolumeType":"gp2","SizeInGB":256},"VolumesPerInstance":4}]}},{"InstanceCount":1,"InstanceGroupType":"MASTER","Name":"Primary","InstanceType":"m5.16xlarge","EbsConfiguration":{"EbsBlockDeviceConfigs":[{"VolumeSpecification":{"VolumeType":"gp3","Iops":3000,"SizeInGB":100,"Throughput":125},"VolumesPerInstance":1}]}},{"InstanceCount":5,"InstanceGroupType":"TASK","Name":"Task - 1","InstanceType":"m5.xlarge","EbsConfiguration":{"EbsBlockDeviceConfigs":[{"VolumeSpecification":{"VolumeType":"gp3","Iops":3000,"SizeInGB":100,"Throughput":125},"VolumesPerInstance":1}]}}]' \
 --bootstrap-actions '[{"Args":[],"Name":"pgx","Path":"s3://pgx-terraform/emr-config/pgx_emr_bootstrap.sh"}]' \
 --scale-down-behavior "TERMINATE_AT_TASK_COMPLETION" \
 --ebs-root-volume-size "97" \
 --auto-termination-policy '{"IdleTimeout":3600}' \
 --region "us-east-1" \
 --profile pgx

```

### M5.16XLarge EMR Cluster (Sparklyr/PySpark bootstrap with Python3.12) 

```{bash}

aws emr create-cluster \
 --name "PGx_emR" \
 --log-uri "s3n://aws-logs-$aws_account-us-east-1/elasticmapreduce/" \
 --release-label "emr-6.15.0" \
 --service-role "arn:aws:iam::$aws_account:role/EMR_DefaultRole" \
 --security-configuration "pgx_emr2" \
 --managed-scaling-policy '{"ComputeLimits":{"UnitType":"Instances","MinimumCapacityUnits":2,"MaximumCapacityUnits":20,"MaximumOnDemandCapacityUnits":20,"MaximumCoreCapacityUnits":20}}' \
 --termination-protected \
 --ec2-attributes '{"InstanceProfile":"EMR_EC2_DefaultRole","EmrManagedMasterSecurityGroup":"sg-063a1a27f01fac9b1","EmrManagedSlaveSecurityGroup":"sg-040af67bc708a5425","AdditionalMasterSecurityGroups":[],"AdditionalSlaveSecurityGroups":[],"SubnetId":"subnet-07950d9f2eb4d936d"}' \
 --applications Name=Hadoop Name=Hive Name=Spark \
 --configurations '[{"Classification":"spark-hive-site","Properties":{"hive.metastore.client.factory.class":"com.amazonaws.glue.catalog.metastore.AWSGlueDataCatalogHiveClientFactory"}},{"Classification":"hive-site","Properties":{"hive.metastore.client.factory.class":"com.amazonaws.glue.catalog.metastore.AWSGlueDataCatalogHiveClientFactory"}}]' \
 --instance-groups '[{"InstanceCount":5,"InstanceGroupType":"CORE","Name":"Core","InstanceType":"m5.16xlarge","EbsConfiguration":{"EbsBlockDeviceConfigs":[{"VolumeSpecification":{"VolumeType":"gp2","SizeInGB":256},"VolumesPerInstance":4}]}},{"InstanceCount":1,"InstanceGroupType":"MASTER","Name":"Primary","InstanceType":"m5.16xlarge","EbsConfiguration":{"EbsBlockDeviceConfigs":[{"VolumeSpecification":{"VolumeType":"gp3","Iops":3000,"SizeInGB":100,"Throughput":125},"VolumesPerInstance":1}]}},{"InstanceCount":5,"InstanceGroupType":"TASK","Name":"Task - 1","InstanceType":"m5.16xlarge","EbsConfiguration":{"EbsBlockDeviceConfigs":[{"VolumeSpecification":{"VolumeType":"gp3","Iops":3000,"SizeInGB":100,"Throughput":125},"VolumesPerInstance":1}]}}]' \
 --bootstrap-actions '[{"Args":[],"Name":"pgx","Path":"s3://pgx-repository/emr-config/pgx_ade_python312_bootstrap.sh"}]' \
 --scale-down-behavior "TERMINATE_AT_TASK_COMPLETION" \
 --ebs-root-volume-size "97" \
 --auto-termination-policy '{"IdleTimeout":3600}' \
 --region "us-east-1" \
 --profile pgx

```


### Setup local environment 

For passing commands between R and Bash environments

```{r sync-renv-bash}

library(jsonlite)
library(here)

cwd <- here::here()

Sys.setenv("CWD"=cwd)

```

### Get EMR Cluster Info

```{sh get-emr-cluster-id}

aws emr list-clusters --active --profile pgx | tee ${CWD}/emr/cluster-config/cluster_info.json

```


```{r parse-emr-cluster-id}

cluster_json <- read_json(here("emr","cluster-config","cluster_info.json"))

cluster_id <- as.data.frame(cluster_json[["Clusters"]][[1]][["Id"]])

names(cluster_id)  <- "cluster_id"

write.table(cluster_id, here("emr","cluster-config","cluster_id.csv"), sep = ";", col.names = FALSE, row.names = FALSE)

cluster_id

# Save as environmental variable
Sys.setenv("CLUSTER_ID"=cluster_id)

```

#### Cluster Configuration

```{sh get-emr-cluster-config}

aws emr describe-cluster --cluster-id ${CLUSTER_ID} --profile pgx | tee ${CWD}/emr/cluster-config/cluster_description.json

```


```{r read-emr-cluster-config}


cluster_description_json <- read_json(here("emr","cluster-config","cluster_description.json"))


```

#### Cluster Head Node

```{sh get-emr-head-node}

aws emr list-instances --instance-group-type MASTER --cluster-id ${CLUSTER_ID} --profile pgx | tee ${CWD}/emr/cluster-config/emr_node.json

```



```{r parse-emr-head-node-info}

emr_json <- read_json(here("emr","cluster-config","emr_node.json"))

emr_dns <- emr_json[["Instances"]][[1]][["PublicDnsName"]]

emr_instance_id <- emr_json[["Instances"]][[1]][["Ec2InstanceId"]]

emr_instance_id <- as.data.frame(emr_instance_id)

write.table(emr_instance_id, here("emr","cluster-config","emr_instance_id.csv"), sep = ";", col.names = FALSE, row.names = FALSE)

emr_dns

emr_instance_id

Sys.setenv("EMR_NODE_ID"=emr_instance_id)

```

#### Allocate Elastic IP to AWS EMR Head Node

```{sh allocate-elastic-ip}

aws ec2 associate-address --instance-id ${EMR_NODE_ID} --allocation-id eipalloc-0d2b32dc1f19a9a4a --profile pgx

```

Setup your elastic IP or use EMR head node DNS. Don't forget to add :8787 at the end for RStudio port
Wait for cluster to build... {about 45 mins}
I have this setup for email notfication


### Go to work!


```{r emr-rstudio-ip}

url <- "http://44.221.159.159:8787"

browseURL(url)


```


For failed cluster starts:

```{bash}

aws emr describe-cluster --cluster-id ${CLUSTER_ID} --profile pgx | tee ${CWD}/cluster-config/bad_cluster_info.json

```