---
title: "EC2 Start Spot Instance"
author: "R. Jerome Dixon"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
editor: source
format:
  html:
    toc: true
    toc-depth: 5
    html-math-method: katex
    df-print: paged
    embed-resources: true
    fontsize: "11pt"
    code-fold: true
    code-summary: "Show the code"
    smooth-scroll: true
editor_options:
  chunk_output_type: inline
execute:
  echo: true
  message: false
  warning: false
  eval: true
---

### Setup local environment

For passing commands between R and Bash environments

```{r sync-renv-bash}

library(jsonlite)
library(here)

cwd <- here::here()

Sys.setenv("CWD"=cwd)

```

### Diagnose AWS Profile Issues

Since you have admin access, let's verify the `pgx` profile is configured correctly.

```{bash}

echo "=== AWS Profile Diagnostics ==="

# Check if pgx profile exists
echo "1. Checking AWS profiles:"
aws configure list-profiles

echo ""
echo "2. Current pgx profile configuration:"
aws configure list --profile pgx

echo ""
echo "3. Testing basic permissions with pgx profile:"
if aws sts get-caller-identity --profile pgx --region us-east-1; then
    echo "✅ Profile authentication works"
else
    echo "❌ Profile authentication failed"
fi

echo ""
echo "4. Testing EC2 permissions specifically:"
if aws ec2 describe-regions --profile pgx --region us-east-1 --output table; then
    echo "✅ EC2 permissions work"
else
    echo "❌ EC2 permissions failed"
fi

```

### Check EC2 Instance Profile Issues

The `EC2_Spot` instance profile might not have the required permissions for Elastic IP operations.

```{bash}

echo "=== EC2 Instance Profile Diagnostics ==="

# Check if EC2_Spot instance profile exists and its permissions
echo "1. Checking EC2_Spot instance profile:"
if aws iam get-instance-profile --instance-profile-name EC2_Spot --profile pgx --region us-east-1; then
    echo "✅ EC2_Spot instance profile exists"
    
    # Get the role associated with this instance profile
    ROLE_NAME=$(aws iam get-instance-profile --instance-profile-name EC2_Spot --profile pgx --region us-east-1 --query 'InstanceProfile.Roles[0].RoleName' --output text)
    echo "Associated Role: $ROLE_NAME"
    
    # Check attached policies
    echo ""
    echo "2. Attached managed policies:"
    aws iam list-attached-role-policies --role-name $ROLE_NAME --profile pgx --region us-east-1 --output table
    
    echo ""
    echo "3. Inline policies:"
    aws iam list-role-policies --role-name $ROLE_NAME --profile pgx --region us-east-1 --output table
    
else
    echo "❌ EC2_Spot instance profile not found or not accessible"
fi

```

### Fix EC2 Instance Profile Permissions

If the instance profile lacks Elastic IP permissions, add them:

```{bash}

echo "=== Adding Elastic IP Permissions to EC2_Spot Role ==="

# Get the role name from the instance profile
ROLE_NAME=$(aws iam get-instance-profile --instance-profile-name EC2_Spot --profile pgx --region us-east-1 --query 'InstanceProfile.Roles[0].RoleName' --output text 2>/dev/null)

if [ "$ROLE_NAME" != "None" ] && [ "$ROLE_NAME" != "null" ] && [ "$ROLE_NAME" != "" ]; then
    echo "Adding Elastic IP permissions to role: $ROLE_NAME"
    
    # Create policy document for Elastic IP operations
    POLICY_DOC='{
        "Version": "2012-10-17",
        "Statement": [
            {
                "Effect": "Allow",
                "Action": [
                    "ec2:AssociateAddress",
                    "ec2:DisassociateAddress",
                    "ec2:DescribeAddresses",
                    "ec2:DescribeInstances"
                ],
                "Resource": "*"
            }
        ]
    }'
    
    # Add inline policy to the role
    if aws iam put-role-policy \
        --role-name $ROLE_NAME \
        --policy-name ElasticIPManagement \
        --policy-document "$POLICY_DOC" \
        --profile pgx --region us-east-1; then
        
        echo "✅ Elastic IP permissions added to EC2_Spot role"
        echo "Wait 1-2 minutes for permissions to propagate, then retry the IP association"
    else
        echo "❌ Failed to add permissions. You may need to do this via AWS Console:"
        echo "1. Go to IAM → Roles → $ROLE_NAME"
        echo "2. Add inline policy with EC2 Elastic IP permissions"
    fi
else
    echo "Could not determine role name. Manual fix needed:"
    echo "1. Go to IAM → Roles → [Your EC2_Spot role]"
    echo "2. Add policy with ec2:AssociateAddress, ec2:DisassociateAddress permissions"
fi

```

### Alternative: Use Default Profile

If the `pgx` profile has issues, try using your default admin profile:

```{bash}

echo "=== Testing Default Profile ==="

# Test with default profile (remove --profile pgx)
echo "1. Default profile identity:"
aws sts get-caller-identity --region us-east-1

echo ""
echo "2. Test Elastic IP move with default profile:"
if [ -f "${CWD}/allocation_id.txt" ] && [ -f "${CWD}/new_instance_id.txt" ]; then
    ALLOCATION_ID=$(cat ${CWD}/allocation_id.txt)
    NEW_INSTANCE_ID=$(cat ${CWD}/new_instance_id.txt)
    
    echo "Using default profile to move Elastic IP..."
    echo "Allocation ID: $ALLOCATION_ID"
    echo "New Instance: $NEW_INSTANCE_ID"
    
    if aws ec2 associate-address \
      --instance-id $NEW_INSTANCE_ID \
      --allocation-id $ALLOCATION_ID \
      --allow-reassociation \
      --region us-east-1; then
        
        echo "✅ Elastic IP successfully moved using default profile!"
        
        # Get the public IP for verification
        PUBLIC_IP=$(aws ec2 describe-addresses \
          --allocation-ids $ALLOCATION_ID \
          --region us-east-1 \
          --query 'Addresses[0].PublicIp' --output text)
        
        echo "Your instance is now available at: $PUBLIC_IP"
    else
        echo "❌ Still failed with default profile"
    fi
else
    echo "Missing allocation_id.txt or new_instance_id.txt files"
fi

```

## EC2 Spot Instance Strategy Summary

This document implements a multi-tiered approach for launching EC2 spot instances optimized for pharmacogenomics analysis workloads:

### Instance Specifications
- **Instance Type**: `x2iedn.8xlarge` (high-memory, NVMe SSD optimized)
- **AMI**: `ami-02434e92bf508ded3` (pre-configured analysis environment)
- **Target Use**: Memory-intensive genomic data processing

### Launch Strategies

#### 1. Existing Instance Management
- Check and restart previously created instances
- Cost-effective for resuming interrupted work

#### 2. Single AZ Spot Launch
- Direct spot instance creation in specific availability zone
- Simple approach for immediate capacity needs

#### 3. Multi-AZ Fleet Strategy (Recommended)
- **Availability Zones**: Spans us-east-1a, us-east-1b, and us-east-1f
- **Allocation Strategy**: Capacity-optimized for best availability and pricing
- **Target Capacity**: Exactly 1 instance total across all zones
- **Benefits**:
  - Maximizes chance of successful spot allocation
  - Automatically selects optimal AZ based on current capacity and pricing
  - Reduces interruption risk through intelligent placement
  - Cost optimization through spot market competition

### Key Configuration
- **Security**: IAM instance profile `EC2_Spot` with appropriate permissions
- **Access**: SSH key `mushin_pgx` for secure connection
- **Networking**: Public subnets across multiple AZs for redundancy
- **Market Type**: 100% spot instances for maximum cost savings

This strategy balances cost optimization, availability, and reliability for large-scale genomic analysis workloads.

---

### Check Instances

```{bash}

aws ec2 describe-instances --query 'Reservations[*].Instances[*].InstanceId' --output json --profile pgx --region us-east-1 | tee ${CWD}/ec2-config/ec2_id.json

```

```{r parse-ec2-cluster-id}

ec2_id_json <- read_json(here("ec2-config","ec2_id.json"))

ec2_id <- as.data.frame(ec2_id_json[[1]][[1]])

names(ec2_id)  <- "ec2_id"

write.table(ec2_id, here("ec2-config","ec2_id.csv"), sep = ";", col.names = FALSE, row.names = FALSE)

ec2_id

# Save as environmental variable
Sys.setenv("EC2_ID"=ec2_id)

```

### Discover VPC and Subnet Configuration

**Important**: Always use existing VPC and subnets when possible to minimize configuration changes, maintain consistent security groups, routing tables, and network ACLs. Creating new networking components can introduce connectivity issues and require additional security configuration.

#### List Available VPCs

```{bash}

echo "=== Available VPCs ==="
aws ec2 describe-vpcs \
  --query 'Vpcs[*].{VpcId:VpcId,CidrBlock:CidrBlock,State:State,IsDefault:IsDefault,Tags:Tags[?Key==`Name`].Value|[0]}' \
  --output table \
  --profile pgx --region us-east-1

```

#### List Available Subnets by VPC

```{bash}

echo "=== Available Subnets by VPC ==="
aws ec2 describe-subnets \
  --query 'Subnets[*].{VpcId:VpcId,SubnetId:SubnetId,AvailabilityZone:AvailabilityZone,CidrBlock:CidrBlock,PublicIP:MapPublicIpOnLaunch,State:State,Name:Tags[?Key==`Name`].Value|[0]}' \
  --output table \
  --profile pgx --region us-east-1

```

#### Save Current Subnet Configuration

```{bash}

# Save subnet details to files for reference
aws ec2 describe-subnets \
  --query 'Subnets[*].{SubnetId:SubnetId,VpcId:VpcId,AvailabilityZone:AvailabilityZone,CidrBlock:CidrBlock,PublicIP:MapPublicIpOnLaunch}' \
  --output json \
  --profile pgx --region us-east-1 > ${CWD}/vpc/subnets.json

echo "Subnet configuration saved to vpc/subnets.json"

# Extract just subnet IDs for easy reference
aws ec2 describe-subnets \
  --query 'Subnets[*].SubnetId' \
  --output text \
  --profile pgx --region us-east-1 > ${CWD}/subnet.txt

echo "Subnet IDs saved to subnet.txt"

```

#### Show Security Groups for Reference

```{bash}

echo "=== Security Groups in VPC ==="
aws ec2 describe-security-groups \
  --query 'SecurityGroups[*].{GroupId:GroupId,GroupName:GroupName,VpcId:VpcId,Description:Description}' \
  --output table \
  --profile pgx --region us-east-1

```

### Start Current Instance

```{bash}

aws ec2 start-instances --instance-ids ${EC2_ID} --profile pgx --region us-east-1

```

#### Run Pre-Configured Spot Instance in Different AZ if Current Instance Not Available

```{bash}
aws ec2 describe-spot-price-history \
  --instance-types x2iedn.8xlarge \
  --product-descriptions "Linux/UNIX" \
  --start-time $(date -u +"%Y-%m-%dT%H:%M:%SZ") \
  --region us-east-1

```

Existing preconfigured image: ami-02434e92bf508ded3

### Run in Desired Zone

```{bash}

aws ec2 run-instances \
  --image-id ami-02434e92bf508ded3 \
  --instance-type x2iedn.8xlarge \
  --subnet-id subnet-5de81a53 \
  --security-group-ids sg-0eb0da772c42415dd \
  --iam-instance-profile Name=EC2_Spot \
  --instance-market-options 'MarketType=spot, SpotOptions={SpotInstanceType=persistent, InstanceInterruptionBehavior=stop}' \
  --key-name mushin_pgx

```

### Run in Available Zone Based on Multiple AZs

```{bash}

LT_JSON=$(cat <<'JSON'
{
  "ImageId": "ami-02434e92bf508ded3",
  "InstanceType": "x2iedn.8xlarge",
  "SecurityGroupIds": ["sg-0eb0da772c42415dd"],
  "IamInstanceProfile": {"Name": "EC2_Spot"},
  "KeyName": "mushin_pgx",
  "InstanceMarketOptions": {
    "MarketType": "spot",
    "SpotOptions": {
      "SpotInstanceType": "persistent",
      "InstanceInterruptionBehavior": "stop"
    }
  }
}
JSON
)

# Create a lightweight Launch Template on the fly
LT_ID=$(aws ec2 create-launch-template \
  --launch-template-name spot-x2iedn-ad-hoc \
  --version-description base \
  --launch-template-data "$LT_JSON" \
  --query "LaunchTemplate.LaunchTemplateId" --output text)

# Use an Instant Fleet with capacity-optimized across all available subnets
SUBNET_A=subnet-5bfc3416   # us-east-1a public
SUBNET_B=subnet-95cecbc9   # us-east-1b public (current)
SUBNET_F=subnet-5de81a53   # us-east-1f public

aws ec2 create-fleet \
  --type maintain \
  --launch-template-configs "LaunchTemplateSpecification={LaunchTemplateId=$LT_ID,Version=\$Latest},Overrides=[{SubnetId=$SUBNET_A},{SubnetId=$SUBNET_B},{SubnetId=$SUBNET_F}]" \
  --spot-options 'AllocationStrategy=capacity-optimized,InstanceInterruptionBehavior=stop' \
  --target-capacity-specification 'TotalTargetCapacity=1,OnDemandTargetCapacity=0,SpotTargetCapacity=1,DefaultTargetCapacityType=spot'

```

### Go to AWS Console - Associate Elastic IP

- Log In
- Sync latest code base