---
title: "Route 53 Setup"
author: "Jerome Dixon"
format: html
editor: visual
---

## Prerequisites

Load environment variables before running commands:

```{bash}
source ../env/load_env.sh
```

## Overview
This document outlines the process of setting up HTTPS access for a static website hosted in an AWS S3 bucket using Route 53 for DNS management and CloudFront for content delivery and SSL termination.

## Step 1: Verify S3 Bucket Contents
First, let's verify the contents of our S3 bucket that contains the static website files.

```{bash}
aws s3 ls jerome-dixon.io/uva/notebooks --recursive
```

```{r}
# Original S3 URL before setting up custom domain
original_url <- 'https://s3.us-east-1.amazonaws.com/jerome-dixon.io/uva/notebooks/lasso_scorecard_model.html'
browseURL(original_url)
```

## Step 2: Route 53 Hosted Zone Setup
We need to identify our Route 53 hosted zone to manage DNS records for our domain.

```{bash}
aws route53 list-hosted-zones-by-name --dns-name jerome-dixon.io
```

Note: When creating alias records, they must use the AWS S3 hosted zone ID as the target.

```{bash eval=FALSE}
# Command to update DNS records using a JSON configuration file
aws route53 change-resource-record-sets \
  --hosted-zone-id Z0675623P625MT2KJFTK \
  --change-batch file://alias_record.json 
```

## Step 3: SSL Certificate Setup
To enable HTTPS, we need to request an SSL certificate from AWS Certificate Manager (ACM).

```{bash eval=FALSE}
# Request a new SSL certificate for our domain
aws acm request-certificate \
  --domain-name jerome-dixon.io \
  --validation-method DNS \
  --region us-east-1 \
  --idempotency-token jeromedixoncert \
  --output json
```

After requesting the certificate, we need to validate domain ownership by adding DNS records.

```{bash}
# Check the validation records needed for the certificate
aws acm describe-certificate \
  --certificate-arn arn:aws:acm:${AWS_REGION_PRIMARY}:${AWS_ACCOUNT_ID_PRIMARY}:certificate/${ACM_CERTIFICATE_ID} \
  --region ${AWS_REGION_PRIMARY} \
  --query "Certificate.DomainValidationOptions[0].ResourceRecord"
```

```{bash eval=FALSE}
# Add the validation record to Route 53
aws route53 change-resource-record-sets \
  --hosted-zone-id Z0675623P625MT2KJFTK \
  --change-batch '{
    "Changes": [
      {
        "Action": "UPSERT",
        "ResourceRecordSet": {
          "Name": "_82acef843a5a9b3f1e4a8eedcb0ef9b9.jerome-dixon.io.",
          "Type": "CNAME",
          "TTL": 300,
          "ResourceRecords": [{ "Value": "_afb1e85eb191b5a7ed7f58e44038a6a9.xlfgrmvvlj.acm-validations.aws." }]
        }
      }
    ]
  }'
```

## Step 4: CloudFront Distribution Setup
Create a CloudFront distribution to serve our S3 content with HTTPS.

```{bash eval=FALSE}
# Create a CloudFront distribution with our SSL certificate
aws cloudfront create-distribution \
  --distribution-config '{
    "CallerReference": "jerome-dixon-https-2025",
    "Aliases": {
      "Quantity": 1,
      "Items": ["jerome-dixon.io"]
    },
    "Origins": {
      "Quantity": 1,
      "Items": [
        {
          "Id": "S3-jerome-dixon.io",
          "DomainName": "jerome-dixon.io.s3-website-us-east-1.amazonaws.com",
          "CustomOriginConfig": {
            "HTTPPort": 80,
            "HTTPSPort": 443,
            "OriginProtocolPolicy": "http-only",
            "OriginSslProtocols": { "Quantity": 1, "Items": ["TLSv1.2"] }
          }
        }
      ]
    },
    "DefaultCacheBehavior": {
      "TargetOriginId": "S3-jerome-dixon.io",
      "ViewerProtocolPolicy": "redirect-to-https",
      "AllowedMethods": {
        "Quantity": 2,
        "Items": ["GET", "HEAD"],
        "CachedMethods": {
          "Quantity": 2,
          "Items": ["GET", "HEAD"]
        }
      },
      "ForwardedValues": {
        "QueryString": false,
        "Cookies": { "Forward": "none" }
      },
      "MinTTL": 0
    },
    "Comment": "jerome-dixon.io HTTPS website",
    "Enabled": true,
    "ViewerCertificate": {
      "ACMCertificateArn": "arn:aws:acm:${AWS_REGION_PRIMARY}:${AWS_ACCOUNT_ID_PRIMARY}:certificate/${ACM_CERTIFICATE_ID}",
      "SSLSupportMethod": "sni-only",
      "MinimumProtocolVersion": "TLSv1.2_2021"
    }
  }'
```

## Step 5: Update Route 53 with CloudFront Distribution
Finally, update Route 53 to point to our CloudFront distribution.

```{bash}
# Update the A record to point to CloudFront
aws route53 change-resource-record-sets \
  --hosted-zone-id Z0675623P625MT2KJFTK \
  --change-batch '{
    "Changes": [
      {
        "Action": "UPSERT",
        "ResourceRecordSet": {
          "Name": "jerome-dixon.io.",
          "Type": "A",
          "AliasTarget": {
            "HostedZoneId": "Z2FDTNDATAQYW2",
            "DNSName": "d2ydx9ea92vqb8.cloudfront.net.",
            "EvaluateTargetHealth": false
          }
        }
      }
    ]
  }'
```

## Verification
Let's verify our setup by checking the DNS records and testing the HTTPS connection.

```{bash}
# Check the A record in Route 53
aws route53 list-resource-record-sets \
  --hosted-zone-id Z0675623P625MT2KJFTK \
  --query "ResourceRecordSets[?Name == 'jerome-dixon.io.']"
```

```{bash}
# Test HTTPS connection
curl -I https://jerome-dixon.io
```

```{r}
# Test the final HTTPS URL
https_url <- 'https://jerome-dixon.io/uva/notebooks/lasso_scorecard_model.html'
browseURL(https_url)
```

