---
title: "emr_gpu"
format: html
editor: visual
---

```{r sync-renv-bash}

library(here)

cwd <- here::here()

Sys.setenv("CWD"=cwd)

```

### AWS EC2 GPU Instance Types for EMR

This section provides a comparison between g4dn, g5, p4d, and p4de instance types. These are GPU-optimized instances suitable for machine learning and deep learning workloads. The table below shows key differences in specifications and use cases:

| Feature | g4dn | g5 | p4d | p4de |
|-----------------|--------------|--------------|--------------|--------------|
| GPU Type | NVIDIA T4 | NVIDIA A10G | NVIDIA A100 | NVIDIA A100 |
| GPU Memory | 16 GB | 24 GB | 40 GB | 80 GB |
| vCPUs (2xlarge) | 8 | 8 | 96 | 96 |
| Memory (2xlarge) | 32 GB | 32 GB | 1152 GB | 1152 GB |
| Network Bandwidth | Up to 25 Gbps | Up to 25 Gbps | Up to 400 Gbps | Up to 400 Gbps |
| Use Case | General ML inference, moderate training | ML training, inference, graphics | Large-scale ML training | Large-scale ML training |
| Price-Performance | Good | Better | Premium | Premium |
| EBS Optimized | Yes | Yes | Yes | Yes |

## Use case recommendations based on GPU instance type characteristics:

**Use g4dn if:** - You need cost-effective GPU instances for ML inference - Your workloads are moderate in size - You need a good balance of CPU and GPU resources - Your applications don't require the latest GPU architecture

**Use g5 if:** - You need better performance than g4dn but don't need the highest-end GPUs - You're doing ML training and inference - You need good price-performance ratio - Your applications benefit from NVIDIA A10G GPUs

**Use p4d if:** - You're running large-scale ML training workloads - You need high GPU memory (40GB per GPU) - You require maximum network bandwidth - Your applications need NVIDIA A100 GPUs

**Use p4de if:** - You need the highest GPU memory available (80GB per GPU) - You're running extremely large ML models - Cost is not the primary concern - You need maximum performance for deep learning workloads

<https://aws.amazon.com/ec2/instance-types/>

```{r}

browseURL("https://aws.amazon.com/ec2/instance-types/")

```

### Spot Pricing

```{bash}

aws ec2 describe-spot-price-history \
  --instance-types g4dn.2xlarge g5.2xlarge p4d.2xlarge p4de.24xlarge \
  --product-descriptions "Linux/UNIX" \
  --start-time $(date -u +"%Y-%m-%dT%H:%M:%SZ") \
  --region us-east-1 \
  --query "SpotPriceHistory[*].[InstanceType,AvailabilityZone,SpotPrice]" \
  --output table

```

#### Lowest price Instance Type

```{bash}

instance_type=$(aws ec2 describe-spot-price-history \
  --instance-types g4dn.2xlarge g5.2xlarge p4d.2xlarge p4de.24xlarge \
  --product-descriptions "Linux/UNIX" \
  --start-time $(date -u +"%Y-%m-%dT%H:%M:%SZ") \
  --region us-east-1 \
  --query "sort_by(SpotPriceHistory, &SpotPrice)[0].[InstanceType]" \
  --output text)
  
echo "$instance_type"
echo "$instance_type" > instance.txt

```

#### Lowest Price Availability Zone

```{bash}

instance_id=$(cat instance.txt)

availability_zone=$(aws ec2 describe-spot-price-history \
  --instance-types $instance_id \
  --product-descriptions "Linux/UNIX" \
  --start-time $(date -u +"%Y-%m-%dT%H:%M:%SZ") \
  --region us-east-1 \
  --query "sort_by(SpotPriceHistory, &SpotPrice)[0].[AvailabilityZone]" \
  --output text)
  
echo "$availability_zone"
echo "$availability_zone" > zone.txt

```

#### Required Subnet for Lowest Price Instance in EMR VPC

```{bash}

zone_id=$(cat zone.txt)
emr_vpc_id="vpc-0ea3d9b6c6c64d0af"
  
aws ec2 describe-subnets \
  --filters "Name=availability-zone,Values=$zone_id" "Name=vpc-id,Values=$emr_vpc_id" \
  --query 'Subnets[*].[SubnetId,AvailabilityZone,VpcId]' \
  --output table


```

#### If Subnet Available in EMR VPC

(if not available - emr_vpc_config.qmd)

```{bash}

zone_id=$(cat zone.txt)
emr_vpc_id="vpc-0ea3d9b6c6c64d0af"
  
vpc_subnet_id=$(aws ec2 describe-subnets \
  --filters "Name=availability-zone,Values=$zone_id" "Name=vpc-id,Values=$emr_vpc_id" \
  --query 'Subnets[*].[SubnetId]' \
  --output text)
  
echo "$vpc_subnet_id"
echo "$vpc_subnet_id" > vpc_subnet.txt

```

### Create EMR GPU Cluster

```{bash}

subnet_id=$(cat vpc_subnet.txt)
instance_type=$(cat instance.txt)
aws_account="535362115856"

aws emr create-cluster \
  --name "pgx-analysis-gpu" \
  --release-label "emr-7.8.0" \
  --log-uri "s3://aws-logs-${aws_account}-us-east-1/elasticmapreduce" \
  --applications Name=Hadoop Name=Hive Name=JupyterEnterpriseGateway Name=Livy Name=Spark \
  --service-role "arn:aws:iam::${aws_account}:role/EMR_DefaultRole" \
  --ec2-attributes KeyName=mushin_pgx,InstanceProfile=EMR_EC2_DefaultRole,SubnetId=$subnet_id \
  --instance-groups \
    InstanceGroupType=MASTER,InstanceCount=1,InstanceType=m4.2xlarge \
    InstanceGroupType=CORE,InstanceCount=9,InstanceType=$instance_type \
    InstanceGroupType=TASK,InstanceCount=1,InstanceType=$instance_type \
  --configurations file://emr7_gpu_config.json \
  --bootstrap-actions Path="s3://pgx-repository/emr-config/emr_gpu_bootstrap.sh",Name="pgx",Args=[] \
  --auto-termination-policy IdleTimeout=1800 \
  --region "us-east-1"



```

#### Get EMR Cluster ID

```{sh get-emr-cluster-id}

aws emr list-clusters --active | tee cluster_info.json

```

```{r parse-emr-cluster-id}

library(jsonlite)
library(here)

cluster_json <- read_json(here("cluster_info.json"))

cluster_id <- as.data.frame(cluster_json[["Clusters"]][[1]][["Id"]])
cluster_name <- as.data.frame(cluster_json[["Clusters"]][[1]][["Name"]])

names(cluster_id)  <- "cluster_id"
names(cluster_name)  <- "cluster_name"

write.table(cluster_id, here("cluster_id.txt"), sep = "\n", col.names = FALSE, row.names = FALSE, quote = FALSE)
write.table(cluster_name, here("cluster_name.txt"), sep = "\n", col.names = FALSE, row.names = FALSE, quote = FALSE)

cluster_id
cluster_name

```

#### Cluster Config Setting from Spark RAPIDS CLI

```{bash}

spark_rapids_user_tools emr bootstrap --help

```

```{bash}

spark_rapids_user_tools emr bootstrap pgx-analysis-gpu \
 --key_pair_path "/c/Projects/mushin_pgx.pem" \
 --dry_run=False \
 --verbose=True


```

```{bash}

aws emr list-clusters --active --query 'Clusters[?Name==`pgx-analysis-gpu`]' --output json

```

#### RAPIDS CLI TOOL (issues on Windows - need to on EMR Cluster directly)

```{bash}

spark_rapids profiling --help


```

```{python}

import boto3
import json

cluster_id = "j-37NO951BC0GX9"
region = "us-east-1"

emr_client = boto3.client("emr", region_name=region)
ec2_client = boto3.client("ec2", region_name=region)

# Detect instance collection type
cluster_desc = emr_client.describe_cluster(ClusterId=cluster_id)
collection_type = cluster_desc["Cluster"].get("InstanceCollectionType", "INSTANCE_GROUP")

instance_types = set()

if collection_type == "INSTANCE_FLEET":
    fleets = emr_client.list_instance_fleets(ClusterId=cluster_id)["InstanceFleets"]
    for fleet in fleets:
        for spec in fleet.get("InstanceTypeSpecifications", []):
            instance_types.add(spec["InstanceType"])
else:
    instances = emr_client.list_instances(ClusterId=cluster_id)["Instances"]
    for inst in instances:
        instance_types.add(inst["InstanceType"])

instance_description = {}
for itype in instance_types:
    response = ec2_client.describe_instance_types(InstanceTypes=[itype])
    itype_data = response["InstanceTypes"][0]
    vcpus = itype_data["VCpuInfo"]["DefaultVCpus"]
    mem_gb = round(itype_data["MemoryInfo"]["SizeInMiB"] / 1024)
    gpu_count = itype_data.get("GpuInfo", {}).get("Gpus", [{}])[0].get("Count", 0)

    instance_description[itype] = {
        "vCPUs": vcpus,
        "memory": mem_gb,
        "gpu": gpu_count,
        "name": itype,
        "category": "custom"
    }

with open("instance_description.json", "w") as f:
    json.dump(instance_description, f, indent=2)

print("instance_description.json created.")

```

```{bash}

spark_rapids profiling \
  --cluster j-37NO951BC0GX9 \
  --platform emr \
  --eventlogs s3://aws-logs-535362115856-us-east-1/elasticmapreduce/j-37NO951BC0GX9 \
  --tools_jar spark-rapids-tools//rapids-4-spark-tools_2.12-25.02.2.jar \
  --output_folder file://cluster_profile



```

```{bash}

cluster_id=$(cat cluster_id.txt)

spark_rapids profiling s3://aws-logs-535362115856-us-east-1/elasticmapreduce/$cluster_id

```

```{bash}

eventlogs-analyzer \
  --s3-log-uri s3://aws-logs-535362115856-us-east-1/elasticmapreduce/ \
  --output-dir ./rapids_analysis_output \
  --start-date 2025-05-01 \
  --end-date 2025-05-03

```

```{sh get-emr-cluster-config}

export CLUSTER_PROPS_FILE=cluster-props.json
cluster_name=$(cat cluster_name.txt)
cluster_id=$(cat cluster_id.txt)

# Get the cluster ID using interpolated $cluster_name
aws emr describe-cluster \
  --cluster-id  $(aws emr list-clusters --query "Clusters[?Name=='$cluster_name'].Id" --output text) \
  > "$CLUSTER_PROPS_FILE"

# Run RAPIDS profiling
spark_rapids_user_tools emr diagnostic \
  --gpu_cluster "$CLUSTER_PROPS_FILE"


```

```{bash}

cluster_name=$(cat cluster_id.txt)

spark_rapids_user_tools emr bootstrap j-F6BX4DQPWATU \
  --key_pair_path "/c/Projects/mushin_pgx.pem" \
  --dry_run=False \
  --verbose=True


```

``` text

%%configure -f
{
  "conf": {
    "spark.executor.cores": "4",
    "spark.executor.memory": "16g",
    "spark.executor.memoryOverhead": "4g",
    "spark.task.resource.gpu.amount": "1",
    "spark.executor.resource.gpu.amount": "1",
    "spark.rapids.sql.concurrentGpuTasks": "2",
    "spark.rapids.memory.pinnedPool.size": "2g",
    "spark.sql.files.maxPartitionBytes": "512m",
    "spark.dynamicAllocation.enabled": "true",
    "spark.task.cpus": "1",
    "spark.yarn.am.memory": "4g",
    "spark.driver.memory": "32g",
    "spark.yarn.executor.memoryOverhead": "4g",
    "spark.jars.packages.resolve.transitive": "true",
    "spark.executor.extraJavaOptions": "--add-exports java.base/sun.net.util=ALL-UNNAMED",
    "spark.driver.extraJavaOptions": "--add-exports java.base/sun.net.util=ALL-UNNAMED",
    "spark.network.timeout": "600s",
    "spark.rpc.askTimeout": "600s"
  }
}
```
